 % 2020 July 26 6.22 pm  spell done 


 % 2018 pesquera

 \documentstyle[12 pt]{article}
 \addtolength{\oddsidemargin}{-1.1in}
 %\addtolength{\textwidth}{1.0in}
 \addtolength{\textwidth}{1.9in}
 \addtolength{\topmargin}{-1.2 in}
 \addtolength{\textheight}{2.5in}
 %\parskip 1 pt

 \newcommand{\newthmwithin}[3]{\newtheorem{#1q}{#2}[#3]
                         \newenvironment{#1}{\begin{#1q}\sf}{\end{#1q}}}

 \newcommand{\newthm}[3]{\newtheorem{#1q}[#2q]{#3}
                         \newenvironment{#1}{\begin{#1q}\sf}{\end{#1q}}}
 \newcommand{\newthmm}[3]{\newtheorem{#1q}[#2q]{#3}
                         \newenvironment{#1}{\begin{#1q}\rm}}

 \newtheorem{theorem}{Theorem}
 %\newtheorem{theorem}{Theorem}[section]
 \newtheorem{lemma}{Lemma}
 \newtheorem{corollary}{Corollary}
 \newtheorem{fact}{Fact}[section]
 \newcommand{\makenewheading}[1]{\begin{tabbing} {\bf #1:} \end{tabbing}}
 \newcommand{\set}[1]{\{ #1 \}}

 \newcommand{\cp}[1]{\mbox{\sc cap}(#1)}
 \newcommand{\dem}[1]{\mbox{\sc dem}(#1)}

 %\newenvironment{proof}{{\it Proof:}}
 \newenvironment{proof}{{\it Proof:}}{$\Box$}
 \newenvironment{sketchproof}{{\it Sketch of Proof:}}{$\Box$}

 \newcommand{\order}[1]{$\mbox{O}\left(#1\right)$}                   

 \newcommand{\rootm}{\sqrt{m}}

 \newcommand{\floorz}{$ (\lfloor|Z|/2  \rfloor -1) $}
 \newcommand{\mfloorz}{ (\lfloor|Z|/2  \rfloor -1)}
 \newcommand{\mceilz}{ (\lceil|Z|/2 \rceil + 1)}
 \newcommand{\ceilz}{$ (\lceil|Z|/2 \rceil ) $}

 \newcommand{\floor}[1]{\lfloor #1 \rfloor}
 \newcommand{\lbound}[1]{$\Omega\left(#1\right)$}

 \newcommand{\mod}[1]{\left| #1 \right|}
 \newcommand{\belongs}{$\in$}

 \def\bed{\begin{description}}
 \def\ennd{\end{description}}
 \def\bee{\begin{enumerate}}
 \def\ene{\end{enumerate}}


 \newcommand{\co}[1]{Corollary \ref{#1}}
 \newcommand{\th}[1]{Theorem \ref{#1}}
 \newcommand{\lem}[1]{Lemma \ref{#1}}
 \newcommand{\eq}[1]{(\ref{#1})}
 \newcommand{\ep}[1]{Equation (\ref{#1})}
 \newcommand{\underx}[1]{\underbrace{~ {#1} ~}}











 \begin{document}
 \thispagestyle{empty}
 \large
 \normalsize
 % \small
 \baselineskip = 1.6 \normalbaselineskip

 \parskip 5 pt
 %\begin{center}
 $.$
 \small
 \bigskip
 \noindent
 {\bf       Talk from Daddy to Robert}, $~~~$  
  July 26, 2020. This statement expresses my sentiments, and I wanted to write them  down.
 % explicitly,
 to avoid any ambiguity.


  % 266

 \large
  \baselineskip =1.35 \normalbaselineskip
 \parskip 4 pt






Many people travel the World during their young 20's
(Even your  father did some of that!  )

Because it is for most people the FREE-st Time in their lives.
I understand why  you are tempted to go to Turkey.

But I will add a few notes to yesterday's talk before starting lesson

I already promised to give you 2K for every month you decide to stay home.

It is NOT RATIONAL to travel abroad in midst of a world-wide
epidemic --- an opinion of both your mother and me.

You are offering to be cheap labor for a country (Turkey) that is a
cheap labor supply for Germany (and which has surplus of refugees
from the Syrian Civil-Religious War).

Airplane Travel ``could'' be again BLOCKED between Turkey and USA.

And while the above sentence with its word ``could''  is speculative,
IT IS NOT A SPECULATIVE POINT that most Ivy League schools will be
3/4 closed in their  Fall Semesters
(If they open ???? EVEN PARTIALLY ? )

In contrast, I could teach you a real logic course in 3 months time,
and you could in future teach logic in philosophy departments.

Let me close this one page note with an ADDED PERSPECTIVE.
It turns out that almost all of the
most recognized
mathematicians and logicians were also textbook writers.  The
most famous textbook writer of all times was probably Euclid.

No one knows what  Euclid did besides writing a geometry textbook
that was quoted over the ages. You and I could do the same in Logic
with a Willard-and-Willard textbook..... And everyone would want
to hire you then as the field expert.

The point is your language abilities and my math knowledge would
make this textbook FLY HIGH AND CLEAR.



 \end{document}

 
\newpage

OLD TEXT

No wife, No responsibilities, No children, No Job, No Mortgage.
Who cares if there is  little money?

Able to live cheaply in a context where few older people can
similarly  do.

Able to eat cheaply. Can of baked beans can be quite satisfying.

JFK, for example, went to Europe before WW2 broke out and returned
from France to England  LITERALLY  on the EXACT day the war started.

Many rich aristocrats did similar (e.g. FDR when he was young)

Willing to stay in crude hostels and camp sites AND ROUGH IT.

When I was in my mid-20's, I used to sleep in my car  near beaches.
When I was 30 and had my first job, I still traveled to two California
conferences and slept in a RENTED CAR.

I didn't travel to EUROPE via bicycle, such as the Super-RICH in their 20's
but I did take two 11-day tours with a supervised teenage group in New England
when I was 15-16 (camping in cramped hostels).

A famous logician called Brower, who lived in Holland, used to take his
notebook and take a 90-day walk through France every summer.  One year,
he lost his notebook in the middle of his walk, and then constructed
an entirely new theory,  WHICH WAS HIS GREATEST LIFE-TIME ACHIEVEMENT.

Your mother also likes wondering around (more than most women).

The UNUSUAL  ASPECT of the current time is that experts about Covid-19,
such as Dr. Faucci, are predicting the next six months will be a challenging
time, with the disease exponentially growing in size, before a vaccine is
available and MASS PRODUCED for the public to use it.

Instead of hundreds of thousands of people dying, it is going to be possibly
Hundreds of millions and possibly LITERALLY BILLIONS of people dying.

Also, there could be an epidemic of violence, robbery and possibly civil
war in impoverished countries.... EVEN IN USA TO LESSER EXTENT.

The closest event to the the coming six months is the start of WW2.
Your grand-mother was on the last train from Prague to Amsterdam before
Jews were forbidden.  Her cousin was on next train, watched Nazis beat her
father to death when he argued with them, and was sterilized and forced
into prostitution to survive WW2 barely. (JFK left France on EXACTLY the
day War broke out.)

It may sound like a selfish,  foolish  remark from a 72-year old man, 
but to go to Turkey before this 6-month wait for the Covid vaccine is
over could be a fatal mistake, similar to the Last Train from Amsterdam.

And you, perhaps with or without me, could write a book, similar to the
Diary of Anne Frank, about your experiences during the Epidemic and
possibly coming Civil War, during the next six months.  At THE VERY LEAST,
you can have your friend Kira publish it, and probably you can go to a
commercial publishing house and maybe even make it into a movie.

You are in an unusual  situation because at age age 22, you have a 72-year
old father and 61-year old mother.... And you need to think about how to
survive the next 6 months before a vaccine arrives and while riots and
horrible theft and stealing could become a second epidemic around
the World.

Let me make a confession  (say in a gentle voice):

When I was 27, my  mother was dying of cancer and I was sadly reluctant
to come home from Harvard when my grandmother lectured me about my
mother dying.

My psychologist then subsequently reinforced her lecture by telling
me that I DID NOT WANT TO admit and RECOGNIZE how horrible events 
were occurring.  He told me my mother ALSO DID NOT WANT TO ADMIT the
reality, BUT SHE HAD NO CHOICE but to face it because events were happening
TOO SUDDENLY .... with her  health deteriorating.

I suspect you, as a young guy feeling your oats and seeing your world
as invulnerable, are also not admitting the reality. The good news is
that it is only a 6-month crisis lying ahead, and they are talking about
a vaccine being ready before December (although its mass manufacturing
might take 2 months longer).

Partially because I was so young, and did not want my mother's illness
to interfere with my life in Boston and search for a job, and my search
for a wife and future mother of my children, I did not want to go home
to visit my mother.  In the end after my grandmother's lecture,
 I started coming home every
4-6 weeks to visit my mother during the last ten months of her life.

The Covid Crisis is only a 6-month interlude before the vaccine becomes
prevalent. You should know that two of my best friends were diagnosed
with cancer during the last four months. (They are Lee Nagel and Jon Silver,
both of whom you know).  Also two other of my firends
had died in the last 20 years,
who you don't know.)

 \baselineskip =1.5 \normalbaselineskip

Debby (my sister) is also dying and unlikely to live 3-5 more years.

I think I am healthy, BUT YOU SHOULD KNOW I have avoided taking a cancer
test THAT EVERYONE OVER THE AGE OF 65  is recommended to take.

I confess I was delinquent in not taking the cancer test, and I am
unable to take it during the next week while you are getting ready to
leave.

My grandmother took the same cancer test at age 72. They found cancer and
removed it successfully. She then lived for 17 more years and never had
cancer again.  BUT I CAN'T have a cancer removed, if I have it (?), without
my family
% being 
present.

And I am not suggesting you avoid travel --- ONLY POSTPONE it until the
Covid vaccine is available.  It is less certain, but a second epidemic
of riots and civil wars 
and violent crimes
is also possible during the next 6 months.

I would be happy to pay you 2K per month, if you don't trust me (?) about
allowing you to postpone your departure until the 6-month Covid waiting period
is over. (You may not want this money, BUT I AM OFFERING it to show you how
seriously I am about considering a minor postponement  ( I will
put my money behind my excessively loud and cranky voice
and mouth.)

Also, if you don't come with me to a minor 100 dollar visit to lawyer to sign
the needed documents, you might forfeit the entire Willard inheritance.
(You need to hear legal details to understand this issue.)

Josh Silver  (who is  45) is trying to persuade his father Jon to move
to Brooklyn. Jon has cancer, and Josh wants Jon to live nearby Josh
in Brooklyn. (I am not even asking you to necessarily live at 14 Grafton
Road during the most troubling crisis since WW2. BUT to be prudently
cautious...,   for you to at least  consider postpone going  quickly abroad  
during
 this
dangerous
6-month period ! )

\newpage

{\bf Added Comments:}

Doctors have recently speculated that even
young children
  who do not
come with Covid symptons can possibly experience brain damage
and other harm to their body.

Don't forget to talk about NP.  (Good guessing is the subtle cure point) 

Talk about Brower.

Jon Silver predicted that instead   of you spending 90 minutes in church,
you would go for the entire wedding PLUS MORE because you can't say no
to your mother (who will persuade you to stay longer).


\end{document}



