1empire Health Card 890 100 573

pamphlet called ``Welcpme to EBD''

     NYSHIP New York State Health Insurance Program
     talk to Civil Service Office

Empire phone 877 7 NYSHIP

civil service phone = 518 457 5754

social security tel = 800 772 1213

NYS december   2020 to Rhannon DiMatinp

NYS december   2020 to ???? 

civil service phone = 518 457 5754

CALL 9 am EXACTLY to get fastest service

press 1 English 
      2 retiree
      1  billing
      1 additional assistance about cost of health insurance 
Don,t press 2   2 how make payment
      1 additional help by human
          estimated wait time less than 5 minutes when call at 9am 
          They still ask if I requested call-back in context of 5minute wait
          at time of 9.09am 9.21am. 
    (On dec 8, I called at 10.25am to get info about who check is payable to)

Info:

Who am I talking to: Misty

      Monthly Premium  568.64
sick Leave Credit = 347.64  
 
2022 medicare 170.10

Net  = 50.90 per month

Annual 12X above = 610.80


note indicate multiple months of premiums

balance is currently 48.36 which I want to continue.
Indicate that in my correspondence.

always contact office the  Monday after first Friday in December

mail letter to 

New York State Department of Civil Service
Employee Benefits Division
PO Box 645 475
Cincinatti, Ohio 45264 - 5475 (dash should be here)

Can mail regularly or however you feel comfortable

Don't Mail to Albany, NY


(Any way you wish to send it. including Express Mail)

To Whom It May Concern

This letter is from Dan E Willard whose empire insurace  ID  = 
and whose billing id =         
Enclosed is a check for xyz to cover all 12 months of my
2022 health insurance  premiums of xyz per month. For safety sake,
please also keep in you possession my balance of 48.36
from prior years. Sincerely Yours 

QUESTIONS 1) WHO IS CHECK PAYABLE TO????  Need to call on Tuesday?
          2) how should    mail (express mail, registered mail?) 
          3) Three month rule?
          4) Check amount
          5) Robert extension?
Questions

0) Contact Retiree Unit for more details and Misty will do
   that right now. Call them 9am 

1) At 65, MY wife must be enrolled Parts A and B of Medicare

2) Always stay away from Part D Medicare

3) When my wife is 65, ask Retiree unit how to handle it if she is not
enrolled in Medicare

4) If i have at least one dependent, NO CHANGE in my premium

5) Beyond 26, i can purchase extra converage for my son
              until age ???? (unknown). Contact 
         Cobra Application for 36 month of coverages
             (2022 rate = 938 per month. If cancelled cannot
              be restored).


6) If I pass and have family coverage, NYSHIP offers 3 free months
   of extended coverage to beneficiaries, but after that they receive
   information about how to do their own seperate coverage.

7) xyz is standard medicare deduction. And Misty never saw anything
   different this year.

8) Wife can enroll in Medicare without getting Social Security by
   NOT taking money out of Social Security by GIVING BANK ACCOUNT
   for DIRECT deductions.

9)  To find providers outside NYS, click on
        empireplanproviders.com
           enter zip code information



i have to find out who check is payable to
abd ALSO contact social security office
or call 800 772 1213 to discover maximal
medicare payment.


They told me 35 days for reissuing cards in alpabetical order

8.59 am called

talked to Morgan in albany 

check payable to NYS Employee Insurance Pending Account

memo should billing ID = 100 111 8610 Just number on memoline

That number SAME EACH YEAR   
xxx

